\Large \bf New Static Analysis Techniques to Detect Entropy Failure Vulnerabilities in Modern Software Projects

### 2018-12-14 paper/paper.tex
more life -drake -michael scott

\begin{abstract}
    In this work we examine the problem of entropy misuse vulnerabilities in
    security-critical software. Our work approaches the problem from an
    ``ounce of prevention is worth a pound of cure'' perspective; by introducing
    and applying a novel static analysis technique \emph{differential taint analysis},
    we propose a method for catching these bugs. Our approach
    leverages the widespread practice of maintaining software version history for
    modern software projects in order to
     provide augmented heuristics to standard static analysis tools to prove the correctness of
     subsequent versions of programs relative to the correctness of an original version.
     Our contributions include the introduction of differential taint analysis
     as a novel approach to this problem domain, as well as a construction to solve
     the differential taint analysis problem using a reduction to standard static
     analysis techniques.
\end{abstract}

@@

\subsection{Debian OpenSSL}

\cite{schneier2008}

\subsection{FreeBSD}

\cite{hovav2016}

@@

Detecting this class of bug often requires an examination of the source code; 
bugs that lead to completely deterministic outputs
 from cryptographic algorithms can often be caught easily during the manual
 testing of code before it is released. Issues which persist past release
 typically result from low-entropy inputs rather than no-entropy inputs.
 As such, these bugs can be difficult to discover and track down. One solution
 to this problem, manual auditing, does work (historically, a manual
 inspection of source code is how these vulnerabilities are disclosed and fixed).
 However effective, manual audits can be an expensive solution, and altogether 
 infeasible
 for many projects to do often, especially those reliant on community donations and gratis
 developer work. On the other hand, static analysis tools have their own issues:
 due to classic hardness results, static analysis algorithms must rely on heuristic
 methods and often cannot provide the sufficient guarantees to developers, even
 if they admit soundness for all of their results.

@@

\section{Preliminaries}

@@

\section{Our Approach}

@@

	These uninterpreted functions are inferred to be pure functions over their variables, so they capture the notion of the taintset at that point. This allows us to assert their equality in the program we pass to CPAChecker.

@@

Although these inference rules are semantically equivalent to the sequential composition, and clearly provide benefit to CPAChecker, this proving strength comes at a cost. Namely, the if statements and the while loops lead to an exponential blowup in the size of the product program being reasoned about as a result of the copied statements.

@@

and

@@

Sequentially composed, we could imagine passing the following program to CPAChecker: Algorithm \ref{alg:seq}.

@@

To evaluate the algorithm laid out in Section \ref{sec:all-together}, we utilized an off-the-shelf state-of-the-art predicate abstraction tool called CPAChecker \cite{beyer2011cpachecker}. This tool was selected because it was the winner in SV-COMP'18, the Competition on Software Verification.

@@

By plugging these product programs into CPAChecker we were able to get the desired verification results. CPAChecker was able to successfully verify the correct changes and successfully fail on the incorrect changes.

Note that our motivating examples were small and simple. Thus, the augmented inference rules evaluated to the same result as the naive inference rules. Our evaluation code snippets were correctly solved by CPAChecker in every variation of our inference rules: the sequential composition, the synchronized composition, and the augmented composition. Further evaluation is needed to confirm or reject our conjectures about the scalability of the various approaches in the design space and to explore the impact of our augmented inferene rules.

@@

Our proposed solution blends taint analysis with relational verification to control the tractability of the safety analysis task. The algorithm proceeds in three primary steps: instrumenting the project, constructing the product program, and inserting assertions/verifying with off-the-shelf static analysis tools. These steps implement two reductions: the instrumentation step converts the original taint analysis problem into a 2-safety property, and the product program construction converts the 2-safety property into a 1-safety property that can be statically verified after insertions are injected in the final step. These constructed product programs were evaluated with an off-the-shelf static analysis tool called CPAChecker.

>>>>

\begin{abstract}
    In this work we examine the problem of entropy misuse vulnerabilities in
    security-critical software. Our work approaches the problem from an
    ``ounce of prevention is worth a pound of cure'' perspective (we do not
    contribute new disclosures of this bug); by introducing
    and applying a novel static analysis technique called \emph{differential taint analysis},
    we propose a method for catching these bugs. Our approach
    leverages the widespread practice of maintaining software version history for
    modern software projects in order to
     provide augmented heuristics to standard static analysis tools to prove the correctness of
     subsequent versions of programs relative to the correctness of an original version.
     Our contributions include the introduction of differential taint analysis
     as a novel approach to this problem domain, as well as a construction to solve
     the differential taint analysis problem using a reduction to standard static
     analysis techniques.
\end{abstract}

@@

\subsection{Debian OpenSSL}
\label{sec:openssl}

The Debian project maintains its own version of OpenSSL. In 2008 a Debian developer
pushed a patch to the Debian OpenSSL project which commented out a few lines of code
marked by Valgrind, a tool
commonly used to identify memory leak bugs. One particular line 
in OpenSSL's random number generator code was one of these lines \cite{schneier2008}:

\begin{center}
\begin{verbatim}
MD_Update(&m,buf,j);
\end{verbatim}
\end{center}

Unfortunately, this line seeded OpenSSL's
RNG with kernel randomness. As a result, the only entropy that the RNG received
was from the process's ID, which is capped at a value of 32,768 on Linux.
This implies that there are only 32,768 possible outputs for the RNG, which
is used by software to generate cryptographic keys. So all keys generated by
this flawed RNG could be easily enumerated by any adversary.

\subsection{FreeBSD}
\label{sec:freebsd}

An entropy failure bug was also found in the FreeBSD project in 2016 \cite{hovav2016}.
This particular vulnerability was a result of the kernel relying on a weak random number
generator (a simple linear congruential generator, well-known to be
vulnerable to basic cryptanalysis) seeded with the current time during bootup. 
After boot the kernel code was supposed to switch to a
source of high-entropy (presumably at this point in time an entropy pool would be
sufficiently full). However, this switch to the secure adaptor was never made, and the kernel continued 
to rely on the weak RNG throughout the lifetime of the system. This bug is particularly
alarming because many userspace applications, including popular security products like 
OpenSSL, rely on the kernel
to generate high-entropy randomness on their behalf. In fact, kernel randomness is
the de-facto recommended source of entropy in the security community.

@@

Detecting this class of bug often requires an examination of the source code; 
bugs that lead to completely deterministic outputs
 from cryptographic algorithms can often be caught easily during the manual
 testing of code before it is released. Issues which persist past release
 typically result from low-entropy inputs rather than no-entropy inputs, such as
 the historical examples discussed in sections \ref{openssl} and \ref{freebsd}.
 As such, these bugs can be difficult to discover and track down. One solution
 to this problem, manual auditing, does work (historically, a manual
 inspection of source code is how these vulnerabilities are disclosed and fixed).
 However effective, manual audits can be an expensive solution, and altogether 
 infeasible
 for many projects to do often, especially those reliant on community donations and gratis
 developer work. On the other hand, static analysis tools have their own issues:
 due to classic hardness results, static analysis algorithms must rely on heuristic
 methods and often cannot provide the sufficient guarantees to developers, even
 if they admit soundness for all of their results.

@@

 In this work, we propose an approach which marries the best of both worlds.
 We introduce a new static analysis technique we call \emph{differential taint analysis}
 which seeks to prove relative or differential correctness between programs (or program versions).
 This allows software projects to only pay for a single audit of their code, and use
 our methods to prove that future updates are relatively secure to this original version.
 Intuitively, proving relational properties between similar programs should be easier;
 in fact, as in other works \cite{barthe2011relational, lahiri2013differential, logozzo2014verification, sousa2016cartesian}
 we reduce the problem to verifying safety properties of a \emph{single} program, and use domain-specific
 properties (taint analysis) to reduce the burden of the verifier.

 \subsection{Outline}

 In section \ref{sec:prelim} we describe the static analysis background necessary
 to understand our results. We then define differential
 taint analysis and the details of our methods of reducing it to standard
 analysis techniques in section \ref{sec:approach}. Finally, we discuss a 
 simple evaluation of our methods in section \ref{sec:eval}.

\section{Preliminaries}
\label{sec:prelim}

@@

\section{Our Approach}
\label{sec:approach}

@@

	These uninterpreted functions are inferred to be pure functions over their variables, so they capture the notion of the taintset at that point. This allows us to assert their equality in the program we pass to CPAchecker.

@@

Although these inference rules are semantically equivalent to the sequential composition, and clearly provide benefit to CPAchecker, this proving strength comes at a cost. Namely, the if statements and the while loops lead to an exponential blowup in the size of the product program being reasoned about as a result of the copied statements.

@@

@@

Sequentially composed, we could imagine passing the following program to CPAchecker: Algorithm \ref{alg:seq}.

@@

To evaluate the algorithm laid out in Section \ref{sec:all-together}, we utilized an off-the-shelf state-of-the-art predicate abstraction tool called CPAchecker \cite{beyer2011cpachecker}. This tool was selected because it was the winner in SV-COMP'18, the Competition on Software Verification.

@@

By plugging these product programs into CPAchecker we were able to get the desired verification results. CPAchecker was able to successfully verify the correct changes and successfully fail on the incorrect changes.

Note that our motivating examples were small and simple. Thus, the augmented inference rules evaluated to the same result as the naive inference rules. Our evaluation code snippets were correctly solved by CPAchecker in every variation of our inference rules: the sequential composition, the synchronized composition, and the augmented composition. Further evaluation is needed to confirm or reject our conjectures about the scalability of the various approaches in the design space and to explore the impact of our augmented inferene rules.

@@

Our proposed solution blends taint analysis with relational verification to control the tractability of the safety analysis task. The algorithm proceeds in three primary steps: instrumenting the project, constructing the product program, and inserting assertions/verifying with off-the-shelf static analysis tools. These steps implement two reductions: the instrumentation step converts the original taint analysis problem into a 2-safety property, and the product program construction converts the 2-safety property into a 1-safety property that can be statically verified after insertions are injected in the final step. These constructed product programs were evaluated with an off-the-shelf static analysis tool called CPAchecker.
